---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

# Import Extra Libraries
library(gt)
library(dplyr)

# Turn Off Scientific Notation
options(scipen = 999)

```

## Xpress Optimization with OMPR

<!-- badges: start -->

<!-- badges: end -->

The `ompr.xpress` package allows the use of the FICO Xpress optimizer to solve mixed integer and linear programming problems with the OMPR package.

## Installation

A FICO Xpress installation and `xpress` package installation are required for `ompr.xpress` to function. The `xpress` package is not available on CRAN, but is rather included with downloads of the Xpress solver from FICO for Xpress verions greater than 8.13.

If a commercial licensed installation is not already available to you, a community version can be download from the link below. Community license is limited to problem with rows + columns \<= 5000.

[FICO® Xpress Community License \| FICO](https://www.fico.com/en/fico-xpress-community-license)

Installation instructions for the `xpress` package are located inside the directory of the Xpress installation. Instructions from FICO are at the link below:

[Installation of the R and Python packages (fico.com)](https://www.fico.com/fico-xpress-optimization/docs/dms2021-04/installguide/dhtml/chapinst1_sec_secrpython.html?scroll=ssecr)

You can install the development version of ompr.xpress from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bmdahl1/ompr.xpress")
```

## Example

We can use a simple police dispatch shift scheduling problem:

A police department must employ a specified number of dispatchers for all 4-hour shifts throughout the day. Each dispatcher must work an 8-hour shift, and shifts from 2000 - 0800 earn a higher wage rate.

The goal of the optimization will be to minimize total wage costs while supplying all needed dispatchers.

```{r}
#| echo: false

# Create GT Table
shift_blank <- data.frame('Shift.Time' = c('0000 - 0400',
                                           '0400 - 0800',
                                           '0800 - 1200',
                                           '1200 - 1600',
                                           '1600 - 2000',
                                           '2000 - 0000'),
                          'Required' = c(6,7,13,13,9,6),
                          'Shift.Rate' = c(45,45,35,35,35,45))

# Show Table
shift_blank |> 
  gt::gt() |> 
  gt::tab_options(table.align = 'left',
                  table.font.size = '10') |> 
  gt::fmt_currency(columns = 'Shift.Rate') |> 
  gt::cols_label(Shift.Rate  = 'Shift Rate') |> 
  as_raw_html()

```

```{r example}
suppressWarnings(suppressMessages(library(ompr.xpress)))
suppressWarnings(suppressMessages(library(ompr)))
suppressWarnings(suppressMessages(library(xpress)))

# Set Shift Parameters
shift_required <-  c(6,7,13,13,9,6)
shift_rate <- c(45, 45, 35, 35, 35, 45)
n_shifts <- 6

# Create OMPR Model
model <- MIPModel() |> 
  add_variable(start_shift[s], type = 'integer', s = 1:n_shifts, lb = 0) |> 
  add_variable(on_shift[s], type = 'continuous', s = 1:n_shifts, lb = 0) |> 
  add_constraint(on_shift[s] == (start_shift[s] + start_shift[s-1]), s = 2:n_shifts) |>
  add_constraint(on_shift[s] == start_shift[s] + start_shift[n_shifts], s = 1) |> 
  add_constraint(on_shift[s] >= shift_required[s], s = 1:n_shifts) |> 
  set_objective(sum_over(on_shift[s]*shift_rate[s], s = 1:n_shifts), sense = 'min')

# Solve with Xpress
model_results <- model |> 
  solve_model(xpress_optimizer())

```

```{r gt_table, echo=FALSE}

# Output to GT Table
shift_schedule <- shift_blank |> 
  mutate(START = model_results$solution[[2]][1:6],
         ON_SHIFT = model_results$solution[[2]][7:12],
         WAGES = ON_SHIFT*(row_number()*shift_rate)) |> 
  relocate(Required, .after = everything())

# Create GT Table
shift_schedule |> 
  gt::gt() |> 
  gt::tab_options(table.align = 'left', 
                  table.font.size = 10) |> 
  # Add Spanner Label
    tab_spanner(
    label = "Decision Variables",
    columns = c(
      START, ON_SHIFT, WAGES
    )
  ) |> 
  gt::cols_label(
    Shift.Time = 'Shift Time',
    START = "Starting",
    ON_SHIFT = "On Shift",
    WAGES = 'Total Hourly Wages'
  ) |> 
  as_raw_html()

```

```{r plot, echo = FALSE, fig.height = 3, fig.width = 6, fig.align = "center"}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "50%"
)

# Plot Required vs. On Shift
plot(x = 1:6, 
     y = shift_schedule$Required, 
     type = 'l', 
     lty = 'solid', 
     col = 'black',
     xlab = 'Shift Number',
     ylab = 'On Shift',
     ylim = c(0, 16),
      cex.main=.75, #change font size of title
     cex.sub=.75, #change font size of subtitle
     cex.lab=.75, #change font size of axis labels
     cex.axis=.75, #change font size of axis text  
     main = 'Shift Schedule - On Shift vs. Required')
lines(x = 1:6, y = shift_schedule$ON_SHIFT, type = 'l', lty = 'dashed', col = 'red')

# Create Legend
legend('topright',
       legend = c('Required','On Shift'),
       col = c('black','red'),
       lty = c('solid','dashed'),
       cex = .75)

```

## Sensitivity Analysis

Sensitivity Analysis will be performed on both LP and MIP problems. When the problem is MIP, then all integer variables will be fixed to the solution from the mixed integer solution, and the problem will be rerun as an LP to perform the sensitivity analysis.

#### Objective Function

Returns upper and lower sensitivity ranges for specified objective function coefficients.\
If the objective coefficients are varied within these ranges the current basis remains optimal and the reduced costs remain valid.\
([objsa (fico.com)](https://www.fico.com/fico-xpress-optimization/docs/latest/solver/optimizer/R/HTML/objsa.html))

```{r sensitivity_analysis_objsa, echo=FALSE}

# Extract Sensitivity Values
model_results$additional_solver_output$obj_sensitivity |> 
  gt::gt() |> 
  gt::tab_options(table.align = 'left', 
                  table.font.size = 10) |> 
  gt::tab_header(title = 'Objective Function Sensitivity Analysis') |> 
  gt::as_raw_html()

```

#### Lower and Upper Bounds

Returns upper and lower sensitivity ranges for specified variables' lower and upper bounds.\
If the bounds are varied within these ranges the current basis remains optimal and feasible.

[bndsa (fico.com)](https://www.fico.com/fico-xpress-optimization/docs/latest/solver/optimizer/R/HTML/bndsa.html)

```{r sensitivity_analysis_bndsa, echo=FALSE}

# Extract Sensitivity Values
model_results$additional_solver_output$bnd_sensitivity |> 
  gt::gt() |> 
  gt::tab_options(table.align = 'left', 
                  table.font.size = 10) |> 
  gt::tab_header(title = 'Upper and Lower Bounds Sensitivity Analysis') |> 
  gt::as_raw_html()


```

#### Right-Hand Side

Returns upper and lower sensitivity ranges for specified right hand side (RHS) function coefficients.\
If the RHS coefficients are varied within these ranges the current basis remains optimal and the reduced costs remain valid.

```{r sensitivity_analysis_rhs, echo=FALSE}

# Create RHS Dataframe
rhs_df <- data.frame(lower = model_results$additional_solver_output$rhs_sensitivity$lower,
                     upper = model_results$additional_solver_output$rhs_sensitivity$upper) |> 
  mutate(constraint = row_number()) |> 
  relocate(constraint, .before = everything())

# # Extract Sensitivity Values
rhs_df |> 
  gt::gt() |> 
  gt::tab_options(table.align = 'left', 
                  table.font.size = 10) |> 
  gt::tab_header(title = 'Righ-Hnd Side Sensitivity Analysis') |> 
  gt::as_raw_html()


```

#### Row Duals

The dual values for the row constraints automatically generated from the LP optimize function.

[getduals (fico.com)](https://www.fico.com/fico-xpress-optimization/docs/latest/solver/optimizer/R/HTML/getduals.html)

```{r sensitivity_analysis_duals, echo=FALSE}

# Extract Sensitivity Values
data.frame(Duals = model_results$additional_solver_output$lp_solution$duals) |> 
  mutate(Constraint = row_number()) |> 
  relocate(Constraint, .before = everything()) |> 
  gt::gt() |> 
  gt::tab_options(table.align = 'left', 
                  table.font.size = 10) |> 
  gt::tab_header(title = 'Row Constraint Dual Values') |> 
  gt::as_raw_html()

```

#### Reduced Costs

The decision variable reduced costs are automatically generated from the LP optimize function.

[calcreducedcosts (fico.com)](https://www.fico.com/fico-xpress-optimization/docs/latest/solver/optimizer/R/HTML/calcreducedcosts.html)

```{r sensitivity_analysis_reduced_cost, echo=FALSE}

# Create Dataframe
reduced_costs <- data.frame(Variable = model_results$solution$Variable,
                            Reduced.Costs = 
                              model_results$additional_solver_output$lp_solution$djs[1:length(model_results$additional_solver_output$lp_solution$djs)-1])

# Create GT
reduced_costs |> 
  gt::gt() |> 
  gt::tab_options(table.align = 'left', 
                  table.font.size = 10) |> 
  gt::cols_label(Reduced.Costs = 'Reduced Costs') |> 
  gt::tab_header(title = 'Decision Variable Reduced Costs') |> 
  gt::as_raw_html()


```
